An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-31 Revision/response.tex

	\caption{Performance of Different Attribute Selectivity (AS) under the Same Algorithm}
	\label{fig:attribute_selectivity}
\end{figure*}
%\begin{figure}[htbp]
%	\centering
%	%	\includegraphics[width=0.8\linewidth]{fig/exp_5_2_1_legend.pdf}
%	\includegraphics[width=\linewidth]{fig/exp_5_1_1_SingleLabel_1thread.pdf}
%	\caption{Different AS under the Same Algorithm}
%	\label{fig:attribute_selectivity_2}
%\end{figure}

%%%%%%%%%%%%%%%%%%%%%% R2 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{metarevision}
		\textit{Run experiments with multi-modal queries (e.g., combining text and images) or more complex query types involving both attribute and range filters.}
	\end{metarevision}
}

\noindent
\textbf{Response:} We believe this is an excellent suggestion. To provide a more comprehensive evaluation of the robustness of each algorithm, we introduce new experiments on two challenging multi-modal datasets.

First, we have incorporated a standard text-to-image dataset, which is widely adopted by existing multi-modal ANN algorithms. In this dataset, the base vectors are images while the query set consists of texts, creating a modality gap that demands stronger algorithm robustness to achieve good performance. Second, to further test performance on complex queries, we have constructed a more challenging dataset with a base set of 500,000 text vectors and 500,000 image vectors, and a query set containing 5,000 text and 5,000 image vectors. On this latter dataset, we perform both attribute filtering and range queries using combined text and image inputs to analyze the algorithm's robustness under more complex, mixed-modality conditions. The experimental results are shown in Figure~\ref{fig:attribute_multimodel} and Figure~\ref{fig:range_multimodel}.\textbf{ (Please see Section 4.1.1, Section 4.2.3 and Secntion 4.3.3, highlighted in blue in the revised manuscript.)  
	%Please refer to the blue highlighted parts in  Section 4.1.1 paragraph 4 and  Section 4.2.4 and Section 4.3.4 under Multi-modal Queries.)
}

As for complex query types that involve both attribute filtering and range queries simultaneously, to the best of our knowledge, existing algorithms—except for general-purpose systems such as Milvus and Faiss—do not support this functionality. For example, ACORN needs to build separate indexes for each scenario and cannot handle attribute filtering and range queries simultaneously. Therefore, we do not consider this type of complex query in our experiments but instead study the two query types separately.

### 2025-07-31 Revision/response.tex

%%%%%%%%%%%%%%%%%%%% Summary %%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{summary}[Summary Comments.]
		\textit{The paper empirically studies performance of hybrid querying methods with respect to two types of queries: attribute filtering (AF) NN search, range filtering (RF) NN search. This is a timely problem with high practical relevance. Nevertheless, several improvements would benefit the paper. We therefore recommend a revision. The main revision points are summarized below, however, we expect that the authors do their best in addressing all points raised in the individual reviews.}
	\end{summary}
}

\noindent

%%%%%%%%%%%%%%%%%%%%%% R1 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{metarevision}
		\textit{Improve the readability of some figures (e.g., performance under different attribute distributions or selectivity).}
	\end{metarevision}
}

\noindent
\textbf{Response:} We sincerely appreciate your valuable suggestions, which provide meaningful improvements to our work. Ensuring the readability of the figures is indeed crucial. We have modified both the figure and the related description in the paper to improve readability and ensure academic rigor. \textbf{(Please see Section 4.2.3, highlighted in blue in the revised manuscript.)}

\textcolor{blue}{"The proximity of the curves in the figure reflects algorithm sensitivity to attribute distributional variations. When the curves converge closely, it demonstrates that the algorithm maintains stable performance when attribute distributions change."}

For performance under different attribute distributions, to illustrate how different algorithms are affected by varying attribute distributions, we added robustness indicators (Stable and Unstable) in the figure. This provides more intuitive visualization of how robust each algorithm performs across different attribute distributions. As shown in Figure\ref{fig:attribute_distribution} below.
%\textbf{Response:} Thank you for your valuable suggestions. We agree that the original figure had readability issues. We have first improved the readability of the attribute selectivity Figure~\ref{fig:attribute_selectivity} by adding comparative legends for each algorithm and standardizing the y-axis scales across different subplots.
%
%
%Furthermore, based on your feedback, we have carefully revised and reanalyzed the experimental sections related to “attribute distribution” and “attribute selectivity” in the paper. 
%
%In real-world applications, attribute distributions are often diverse, such as long-tail, uniform, or normal distributions. In our initial design, we included multiple attribute distributions to evaluate the robustness of algorithms under different data characteristics. However, we later realized that attribute distribution is essentially a mixture of different attribute selectivities. Its impact on algorithm performance is ultimately determined by attribute selectivity in each query.
%For example, in a long-tail distribution, querying a common “head” attribute is equivalent to a high attribute selectivity (High AS) case, while querying a rare “tail” attribute is equivalent to a low attribute selectivity (Low AS) case.
%Therefore, by evaluating performance under a range of attribute selectivities (e.g., AS = 1\% to AS = 75\%), we have already covered the effects of different attribute distributions. In this sense, the attribute selectivity experiments provide a unified and quantitative way to reflect the influence of attribute distributions.
%
%We present the results of attribute distribution and attribute selectivity in Figure~\ref{fig:attribute_distribution} and Figure~\ref{fig:attribute_selectivity} , respectively. The conclusions are highly consistent. For example, Faiss performs poorly under high attribute selectivity and also under long-tail distributions. This is because most queries in a long-tail distribution correspond to high-selectivity attributes, while only a few are low-selectivity, which limits the overall performance.
%
%Moreover, previous studies (e.g., FilteredDiskANN, Milvus) also commonly use attribute selectivity as a key metric for evaluating filtering performance.
%
%Based on these reasons, we decided to remove the attribute distribution experiments and retain only the attribute selectivity analysis. We also removed the original Figure~\ref{fig:attribute_selectivity_2}, which displayed too many algorithms in one plot and was difficult to interpret. We believe these changes improve the clarity and readability of the paper, without reducing the completeness of the evaluation.

@@

\begin{figure*}[t]  % 或者 [b] 放到底部
	\centering
%	\includegraphics[width=\linewidth]{fig/exp_3_1_legend.pdf}
	\includegraphics[width=\linewidth]{fig/exp_3_1.pdf}
	\caption{Effect of Attribute Distribution on Query Performance}
	\label{fig:attribute_distribution}
\end{figure*}

For performance under different attribute selectivity, to reflect the performance differences of each algorithm under different attribute selectivity, we added performance rankings for each algorithm across different selectivity levels in the figure. This provides a clearer illustration of how each algorithm is affected by attribute selectivity rates. As shown in Figure~\ref{fig:attribute_selectivity} below.

We also removed the  Figure~\ref{fig:attribute_selectivity_2}, which displayed too many algorithms in one plot and was difficult to interpret and somewhat redundant with Figure~\ref{fig:attribute_selectivity}. We believe these changes improve the clarity and readability of the paper, without reducing the completeness of the evaluation.

@@

	% 第二行：两张子图，左侧宽度是右侧的两倍
	\begin{subfigure}{0.627\textwidth} % 2/3 宽度
		\centering

		\includegraphics[width=0.96\textwidth]{fig/exp_5_2_2.pdf}
		\caption{Algorithms Optimized for High AS}
		\label{fig:exp_5_2_2}
	\end{subfigure}
	\hspace{1mm} % 增加水平间距
	\begin{subfigure}{0.33\textwidth} % 1/3 宽度
		\centering

		\includegraphics[width=0.96\textwidth]{fig/exp_5_2_3.pdf}
		\caption{Algorithms Insensitive to AS}
		\label{fig:exp_5_2_3}
	\end{subfigure}

	\caption{Different AS under the Same Algorithm}
	\label{fig:attribute_selectivity}
\end{figure*}
\begin{figure}[htbp]
	\centering
	%	\includegraphics[width=0.8\linewidth]{fig/exp_5_2_1_legend.pdf}
	\includegraphics[width=\linewidth]{fig/exp_5_1_1_SingleLabel_1thread.pdf}
	\caption{Different AS under the Same Algorithm}
	\label{fig:attribute_selectivity_2}
\end{figure}

@@

	\caption{Comparison of Real and Synthetic Attributes}
	\label{fig:range_2_2}

\end{figure}

%%%%%%%%%%%%%%%%%%%%%% R4 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{metarevision}
		\textit{Instead of using the point ID, run experiments with meaningful attributes for RF-NN search on datasets (Deep, YT-Audio, WIT).}
	\end{metarevision}
}

\noindent
\textbf{Response:} 
Thank you for your suggestion. Using ID as an attribute is a prevalent method in current range filtering algorithm evaluations, so we also use ID as attributes in our experiments. To conduct a meaningful evaluation, we have updated two of our datasets (YT-Audio-Real, WIT-Real) to include real attributes.  \textbf{(Please see Section 4.1.1, highlighted in blue in the revised manuscript.)}

### 2025-07-31 VLDB2026/main.tex

On Text2image-Mix dataset,as shown in Figure~\ref{fig:attribute-multimodal}, UNG and NHQ consistently perform best, followed by CAPS.

@@

		In contrast, the Text2image dataset shown in Figure~\ref{fig:attribute-multimodal} is different. The query set and the base dataset belong to two completely different modalities. Queries are far from the dataset points, and the ground truths are more dispersed. To find the ground truths, algorithms typically need to examine more points, which leads to degraded performance.

>>>>

On Text2image-Mix dataset, as shown in Figure~\ref{fig:attribute-multimodal}, UNG and NHQ consistently perform best, followed by CAPS.

@@

		In contrast, the Text2image dataset shown in Figure~\ref{fig:attribute-multimodal} is different. The query set and the base dataset belong to two completely different modalities. Queries are far from the dataset points, and the ground truths are more dispersed. To find the groundtruths, algorithms typically need to examine more points, which leads to degraded performance.
